SEER Medicaid

### 2023-12-15 Manuscript Analysis code 0 to 19.Rmd
adding revision changes

        mutate(race_cat2 = case_when(race_cat == 0 ~ 0, # Non-Hispanic White
                                     race_cat == 1 ~ 1, # Non-Hispanic Black
                                     race_cat %in% c(2,3,4,5) ~ 2)) %>% # Non-Hispanic Other

@@

        mutate(can_dead = case_when(SEERcausespecificdeathclassific == 0 ~0, # Alive or dead of other cause
                                SEERcausespecificdeathclassific == 1 ~1)) # Dead from cancer
# note 8 an 9 are missing or N/A

@@

# Factor variables
seer0to39$age_cat<-factor(seer0to39$age_cat, levels=c(0:4), labels=c("0-4", "5-9", "10-14", "15-19", "20-39"))
seer0to39$race_cat<-factor(seer0to39$race_cat, levels=c(0:4), labels=c("White", "Black", "AIAN", "API", "Unknown"))
seer0to39$race_cat2<-factor(seer0to39$race_cat2, levels=c(0:2), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic Other"))
seer0to39$hispanic<-factor(seer0to39$hispanic, levels=c(0:1), labels=c("Non-Hispanic", "Hispanic"))
seer0to39$race_ethnicity<-factor(seer0to39$race_ethnicity, levels=c(0:4), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other",  "Hispanic")) 
seer0to39$sex <- factor(seer0to39$sex, levels = c(0:2), labels = c("Male", "Female", "Not Stated"))

@@

# calculate follow-up time from month of dx, recode year of dx TO Month_of_last_follow_up_recode, Year_of_last_follow_up_recode. The difference between these variable dates would provide months of follow-up
summary(seer0to39$Month_of_diagnosis_recode)
summary(seer0to39$Year_of_diagnosis)
summary(as.numeric(seer0to39$Month_of_last_follow_up_recode))
seer0to39$Month_of_last_follow_up_recode<-as.numeric(seer0to39$Month_of_last_follow_up_recode) # make variable numeric
seer0to39$SEER_DateofDeath_Month<-as.numeric(seer0to39$SEER_DateofDeath_Month) # make variable numeric
summary(seer0to39$Month_of_last_follow_up_recode)
summary(seer0to39$Year_of_last_follow_up_recode)
summary(seer0to39$SEER_DateofDeath_Month)
summary(seer0to39$SEER_DateofDeath_Year)

# assign a date of diagnosis using 15 as the day because there is no day of dx so assume mean is 15
seer0to39$proxy_day_of_dx <-15
seer0to39$proxy_day_of_fu <-15
seer0to39$proxy_day_of_dth <-15

@@

seer0to19b <- seer0to19

@@

# recode a category for not in medicaid
seer0to19 <- seer0to19 %>%
  mutate(mcont66 = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb == 0 ~ 1, # continuous
                             mcontb == 1 ~ 2, # discontinuous
                             mcontb == 2 ~ 3)) %>% # Other
  mutate(mcont66b = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb2 == 0 ~ 1, # continuous
                             mcontb2 == 1 ~ 2, # discontinuous
                             mcontb2 == 2 ~ 3, # Other
                             mcontb2 == 3 ~ 4)) # at or after diagnosis

@@

# create factor variable for analyses
seer0to19$mcont66 <- factor(seer0to19$mcont66, levels = c(0:3), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other"))
seer0to19$mcont66b <- factor(seer0to19$mcont66b, levels = c(0:4), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"))

@@

table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19)

table1(~can_type|In_Medicaid_Flag, seer0to19)

table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66b, droplevels = TRUE, seer0to19)
```

# FOR REVISION (insurance distribution)
```{r}
# classify primary payor variable
class(seer0to19$Primary_Payer_at_DX)
seer0to19$Primary_Payer_at_DX <- as.numeric(seer0to19$Primary_Payer_at_DX)

@@

# the mlogit.cluster function does not like labels
seer0to19_stage <- seer0to19_stage %>% 
  remove_labels()
```

@@

df20<-df20[-c(1, 5:25, 29:48),]
df20 <- tibble::rownames_to_column(df20, "Category") # Apply rownames_to_column
df20  
```

@@

# Mcont66
df3a <- data.frame(matrix(ncol=6, nrow=0)) #create the table
columns<-c("Category", "OR","cluster.lowCI", "cluster.highCI", "CancerType") #label the columns with names
colnames(df3a)<-columns

@@

Plot3 <- SuppTable1a %>%
  filter(`Exposure comparison` != "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Regional vs. in situ/localized") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`,  ymax = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "C. Regional vs. in situ/localized", x = "Cancer type") +
  ylim(0,4.0) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                        values=c("red", "cyan3", "blue"))

Plot3
table(SuppTable1a$`Exposure comparison`)

Plot4 <- SuppTable1a %>%
  filter(`Exposure comparison` != "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Distant vs. in situ/localized") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`,  max = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
 # facet_grid(cols = vars(y.level)) +
  labs(color = NULL, title = "D. Distant vs. in situ/localized", x = "Cancer type") +
  ylim(0,5.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled", "Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                        values=c("red", "cyan3", "blue"))

@@

dev.off()
```
# Survival plots with discontinuous divided to at diagnosis and the rest
```{r}
# mcont66b
 plot3 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ mcont66b, data = seer0to19 ), 
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct", 
    title = "B",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(12, "bold"),
    legend.labs = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"),
    palette = c("red", "cyan3", "blue", "black", "purple"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(50, 100),
    censor.size=0.5,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(2,55),
    conf.int = FALSE)

@@

pdf("Figure 3.pdf", width=8.5, height=11, paper="USr")
plot1$plot + plot2$plot + plot_layout(ncol=1, nrow=2, guides='keep') &
  theme(legend.position='bottom')

# SUPPLEMENTARY TABLE 2: Estimate survival time 0 to 19 at 60 mo
```{r}
# Overall
overall <- summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to19), times=60) # estimates 60 month survival time

overall <- as.data.frame(cbind(overall$time, round(overall$surv*100, 2), round(overall$lower*100, 2), round(overall$upper*100,2))) # makes dataframe
overall$Category <- "Overall"
overall <- setNames(overall, c("Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Category")) # gives column names
overall <- overall %>% select("Category","Time (months)", "Survival Percentage", "Lower CI", "Upper CI")

@@

# adjusted models 0 to 19
res.cox2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox2
summary(res.cox2)

@@

# Test of the ph assumption for adjusted models
```{r}
# adjusted models 0 to 19
Medflag <- cox.zph(res.cox2, terms = FALSE)
Medflag

# mcont66 adjusted models 0 to 19
Med4cat <- cox.zph(res.cox4, terms = FALSE)
Med4cat
# violated for both

@@

# Reordering
seer0to19.cp.format <- seer0to19.cp.format[with(seer0to19.cp.format, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo
summary(seer0to19$fumo_half)

@@

library(multcomp)
X<- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1)

@@

W <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) 
X <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1)
Y <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1)
Z <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1)

x <- coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format) 
  t0<- glht(x, linfct = W) 
  df1<-confint(t0)
  df1 <- as.data.frame(exp(df1$confint))

@@

# Graph hazard ratios and confidence intervals across time for enrolled variable
```{r}
p<- results %>%
  ggplot() +
  geom_point(aes(y=Estimate, x= interval), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() 
p
```
# SUPPLEMENTARY TABLE 6b: Which cancer is driving this? 
```{r}
# In_Medicaid_Flag
results <- data.frame(matrix(ncol=5, nrow=0)) # create empty table for results that will be populated by for loop
columns<-c("estimate", "lwr", "upr", "age", "interval") # label the columns with names
colnames(results) <- columns

@@

# Group by cancer type and plot
```{r}
SuppTable6b %>%
  group_by(Group) %>%
  ggplot() +
  geom_point(aes(y=Estimate, x=interval)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() + 
  facet_wrap(~Group, nrow=4, ncol=2, scales = "free") +
  labs( x = "Time interval") + 
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=12, face="bold"))
```
# run models based on intervals for mcont66
```{r}
# testing here
m <-coxph(SurvObj ~ mcont66 + mcont66*strata(interval) + age_cat + race_ethnicity +  Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format) 
m

X <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1)

@@

A <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # discontinuous 1st interval
B <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # other 1st interval
C <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Not enrolled 1st interval

D <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0), 1) # discontinuous 2nd interval
E <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0), 1) # other 2nd interval
F <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0), 1) # Not enrolled 2nd interval

G <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0), 1) # discontinuous 3rd interval
H <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0), 1) # other 3rd interval
I <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0), 1) # Not enrolled 3rd interval

J <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1) # discontinuous 3rd interval
K <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1) # other 3rd interval
L <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1) # Not enrolled 3rd interval

@@

SuppTable7c <- rbind(a,b,c,d,f,h,i,j,k, m,n,o,p,r,s,t,u,v,w)

>>>>

        mutate(race_cat2 = case_when(race_cat == 0 ~ 0, # Non-Hispanic White
                                     race_cat == 1 ~ 1, # Non-Hispanic Black
                                     race_cat %in% c(2,3,4) ~ 2)) %>% # Non-Hispanic Other

@@

        mutate(can_dead = case_when(SEERcausespecificdeathclassific == 0 ~0, # Alive or dead of other cause
                                SEERcausespecificdeathclassific == 1 ~1)) # Dead from cancer
# note 8 an 9 are dead missing COD and N/A

@@

# Factor variables
seer0to39$age_cat<-factor(seer0to39$age_cat, levels=c(0:4), labels=c("0-4", "5-9", "10-14", "15-19", "20-39"))
seer0to39$race_cat<-factor(seer0to39$race_cat, levels=c(0:4), labels=c("White", "Black", "AIAN", "API", "Unknown"))
seer0to39$race_cat2<-factor(seer0to39$race_cat2, levels=c(0:2), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic Other"))
seer0to39$hispanic<-factor(seer0to39$hispanic, levels=c(0:1), labels=c("Non-Hispanic", "Hispanic"))
seer0to39$race_ethnicity<-factor(seer0to39$race_ethnicity, levels=c(0:4), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other",  "Hispanic")) 
seer0to39$sex <- factor(seer0to39$sex, levels = c(0:2), labels = c("Male", "Female", "Unknown"))

@@

# calculate follow-up time from month of dx, recode year of dx TO Month_of_last_follow_up_recode, Year_of_last_follow_up_recode. The difference between these variable dates provides months of follow-up
summary(seer0to39$Month_of_diagnosis_recode)
summary(seer0to39$Year_of_diagnosis)
summary(as.numeric(seer0to39$Month_of_last_follow_up_recode))
seer0to39$Month_of_last_follow_up_recode <- as.numeric(seer0to39$Month_of_last_follow_up_recode) # make variable numeric
seer0to39$SEER_DateofDeath_Month <- as.numeric(seer0to39$SEER_DateofDeath_Month) # make variable numeric
summary(seer0to39$Month_of_last_follow_up_recode)
summary(seer0to39$Year_of_last_follow_up_recode)
summary(seer0to39$SEER_DateofDeath_Month)
summary(seer0to39$SEER_DateofDeath_Year)

# assign a date of diagnosis using 15 as the day because there is no day of dx so assume mean is 15
seer0to39$proxy_day_of_dx <- 15
seer0to39$proxy_day_of_fu <- 15
seer0to39$proxy_day_of_dth <- 15

@@

seer0to19b <- seer0to19 # using to not write over seer0to19

@@

# recode a category for not in medicaid
seer0to19 <- seer0to19 %>%
  mutate(mcont66 = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb == 0 ~ 1, # continuous
                             mcontb == 1 ~ 2, # discontinuous
                             mcontb == 2 ~ 3)) %>% # Other
  # not used for this paper
  mutate(mcont66b = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb2 == 0 ~ 1, # continuous
                             mcontb2 == 1 ~ 2, # discontinuous
                             mcontb2 == 2 ~ 3, # Other
                             mcontb2 == 3 ~ 4)) # at or after diagnosis

@@

# create factor variable for analyses
seer0to19$mcont66 <- factor(seer0to19$mcont66, levels = c(0:3), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other"))

seer0to19$mcont66b <- factor(seer0to19$mcont66b, levels = c(0:4), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"))

@@

# used in paper
table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19)

# table1(~can_type|In_Medicaid_Flag, seer0to19)

# table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66b, droplevels = TRUE, seer0to19)
```

# FOR REVISION response (insurance distribution)
```{r}
# classify primary payor variable
class(seer0to19$Primary_Payer_at_DX)
seer0to19$Primary_Payer_at_DX <- as.numeric(seer0to19$Primary_Payer_at_DX)

@@

# the mlogit.cluster function does not like labels
seer0to19_stage <- seer0to19_stage %>% 
  remove_labels()

table(seer0to19_stage$mcont66, seer0to19_stage$stage)

@@

df20<-df20[-c(1, 5:25, 29:48),]
df20 <- tibble::rownames_to_column(df20, "Category") # Apply rownames_to_column
df20

p
```

@@

# mcont66
df3a <- data.frame(matrix(ncol=6, nrow=0)) #create the table
columns<-c("Category", "OR","cluster.lowCI", "cluster.highCI", "CancerType") #label the columns with names
colnames(df3a)<-columns

@@

Plot3 <- SuppTable1a %>%
  filter(`Exposure comparison` != "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Regional vs. in situ/localized") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`,  ymax = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "C. Regional vs. in situ/localized", x = "Cancer type") +
  ylim(0,4.0) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous vs. Not enrolled","Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                        values=c("red", "cyan3", "blue"))

Plot3

table(SuppTable1a$`Exposure comparison`)

Plot4 <- SuppTable1a %>%
  filter(`Exposure comparison` != "Enrolled vs. Not enrolled", `Stage outcome comparison` == "Distant vs. in situ/localized") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`,  max = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
 # facet_grid(cols = vars(y.level)) +
  labs(color = NULL, title = "D. Distant vs. in situ/localized", x = "Cancer type") +
  ylim(0,5.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c( "Discontinuous vs. Not enrolled","Continuous vs. Not enrolled", "Other vs. Not enrolled"),
                        values=c("red", "cyan3", "blue"))

@@

dev.off()
```
# Survival plots with discontinuous divided to at diagnosis and the rest (NOT IN PAPER)
```{r}
# mcont66b
 plot3 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ mcont66b, data = seer0to19 ), 
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct", 
    title = "B",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(12, "bold"),
    legend.labs = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"),
    palette = c("red", "cyan3", "blue", "black", "purple"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(50, 100),
    censor.size=0.5,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(2,55),
    conf.int = FALSE)

@@

# SUPPLEMENTARY TABLE 2: Estimate survival time 0 to 19 at 60 mo
```{r}
# Overall
overall <- summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to19), times=60) # estimates 60 month survival time overall

overall <- as.data.frame(cbind(overall$time, round(overall$surv*100, 2), round(overall$lower*100, 2), round(overall$upper*100,2))) # makes dataframe

overall$Category <- "Overall"

overall <- setNames(overall, c("Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Category")) # gives column names

overall <- overall %>% select("Category","Time (months)", "Survival Percentage", "Lower CI", "Upper CI")

@@

# adjusted models 0 to 19
res.cox2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race, age and a measure of poverty
res.cox2
summary(res.cox2)

@@

# Test of the ph assumption for adjusted models
```{r}
# adjusted models 0 to 19
Medflag <- cox.zph(res.cox2, terms = TRUE)
Medflag

# mcont66 adjusted models 0 to 19
Med4cat <- cox.zph(res.cox4, terms = TRUE)
Med4cat
# violated for both

@@

# Reordering
seer0to19.cp.format <- seer0to19.cp.format[with(seer0to19.cp.format, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo

summary(seer0to19$fumo_half)

@@

library(multcomp)
X<- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # number of 0's and 1's in vector is equal to number of coefficients in model

@@

# set up matrices to calculate HRs for different intervals
W <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) 
X <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1)
Y <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1)
Z <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1)

x <- coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format)

  t0<- glht(x, linfct = W) 
  df1<-confint(t0)
  df1 <- as.data.frame(exp(df1$confint))

@@

# Graph hazard ratios and confidence intervals across time for enrolled variable
```{r}
# factor for order
results$interval <- factor(results$interval, levels = c("0 to ≥12", ">12 to ≥24", ">24 to ≥36", "≥36"))

p<- results %>%
  ggplot() +
  geom_point(aes(y=Estimate, x= interval), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() 
p
```
# SUPPLEMENTARY TABLE 6b: Which cancer is driving this? 
```{r}
# In_Medicaid_Flag
results <- data.frame(matrix(ncol=5, nrow=0)) # create empty table for results that will be populated by for loop
columns<-c("estimate", "lwr", "upr", "age", "interval") # label the columns with names
colnames(results) <- columns

@@

# Group by cancer type and plot
```{r}

plot <- SuppTable6b %>%
  group_by(Group) %>%
  ggplot() +
  geom_point(aes(y=Estimate, x=interval)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() + 
  facet_wrap(~Group, nrow=4, ncol=2, scales = "free") +
  labs( x = "Time interval") + 
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=12, face="bold"))

plot
```
# run models based on intervals for mcont66
```{r}
# testing here
m <-coxph(SurvObj ~ mcont66 + mcont66*strata(interval) + age_cat + race_ethnicity +  Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format) 
m

X <- matrix(c(0,1,0,0,0,
              0,0,0,0,0,
              0,0,0,0,0,
              0,0,0,0,0,
              0,0,0), 1) # 23 coefficients

@@

A <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Continuous 1st interval
B <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Discontinuous 1st interval
C <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Other 1st interval

D <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0), 1) # Continuous 2nd interval
E <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0), 1) # Discontinuous 2nd interval
F <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0), 1) # Other 2nd interval

G <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0), 1) # Continuous 3rd interval
H <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0), 1) # Discontinuous 3rd interval
I <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0), 1) # Other 3rd interval

J <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1) # Continuous 4th interval
K <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1) # Discontinuous 4th interval
L <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1) # Other 4th interval

@@

SuppTable7c <- rbind(a,b,c,d,f,h,i,j,k, m,n,o,p,r,s,t,u,v,w)
```
